Relevance-based Quantization of Scattering Features for Unsupervised Mining of Environmental Audio

### 2018-02-19 paper/structureScene.tex
add editors requirements

\author{Vincent Lostanlen\thanks{This work is supported by the ERC InvariantClass grant 320959.}, Gr\'egoire Lafay, Joakim And\'en, and Mathieu Lagrange}

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
The emerging field of computational ecoacoustics aims at retrieving high-level information from acoustic scenes recorded by some network of sensors. These networks gather large amounts of data requiring analysis. To decide which parts to inspect manually, we need tools that automatically mine the data, identifying recurring patterns and isolated events. This requires a similarity measure for acoustic scenes that does not impose strong assumptions on the data.

The state of the art in audio similarity measurement is the ``bag-of-frames'' approach, which models a recording using summary statistics of short-term audio descriptors, such as mel-frequency cepstral coefficients (MFCCs). They successfully characterize static scenes with little variability in auditory content, but cannot accurately capture scenes with a few salient events superimposed over static background.
To overcome this issue, we propose a two-scale representation which describes a recording using clusters of scattering coefficients. The scattering coefficients capture short-scale structure, while the cluster model captures longer time scales, allowing for more accurate characterization of sparse events. Evaluation within the acoustic scene similarity framework demonstrates the superiority of the proposed approach.
\keywords{
unsupervised learning \and data mining \and acoustic signal processing \and wavelet transforms \and audio databases \and content-based retrieval \and nearest neighbor searches \and acoustic sensors \and environmental sensors.}
% \PACS{PACS code1 \and PACS code2 \and more}
\end{abstract}

@@

The amount of audio data recorded from our sonic environment has grown considerably over the past decades.
In order to measure the effect of human activity and climate change on animal biodiversity, researchers have recently undertaken a massive deployment of acoustic sensors throughout the world \cite{warren2006urban, NessSST13, stowell13b}.
In addition, recent work has explored acoustic monitoring for characterization of human pleasantness in urban areas \cite{guyot2005urban, ricciardi2015sound}, as well as the prediction of annoyance due to traffic \cite{gloaguen}.
Since they bear a strong societal impact and raise many scientific challenges, we believe that these applications are of considerable interest for to signal processing community.

These questions are explored in an emerging field of computational bioacoustics, known as ecoacoustics, which is the interdisciplinary study of relationships between natural or anthropogenic sounds and their environments at the intersection of ecology, acoustics and computer science \cite{krause}.
An important problem in this field is that manually analysing the recorded data to identify the quantities of interest is very costly \cite{wimmer2013sampling}. Some sort of pre-screening is therefore required to reduce the need for human expert listening and annotation.

The most straightforward approach is to specify a closed set of sound classes, such as sounds of animals expected to live near the acoustic sensors. We then train models for these classes which are used to automatically annotate our recordings \cite{}. \ja{What's the reference here?} A given time interval (e.g., a single day) is then represented by the number of events detected during that interval for each class. This allows the scientist to drastically reduce the amount of information requiring manual processing. However, this approach has two drawbacks. First, it relies on trained models whose behavior on unseen data (such as different sensors) is prone to errors and cannot be trusted. Second, and more importantly, it is based on \textit{a priori} knowledge and thus cannot be considered for exploratory analysis where quantities of interest have yet to be defined.

To identify which parts need human inspection, we need tools that can detect both recurring patterns and sparsely distributed events. Identifying recurring patterns allows the user to focus on certain time points for manual annotation, while detection of more rare structures enables discovery of unforeseen phenomena.

With this aim, we need to design an algorithm for acoustic similarity retrieval, where the audio fragments judged ``most similar'' to a given query recording must be extracted from some larger dataset. This requires us to represent an audio recording in a way that captures its distinctive qualities. This has previously been attempted using bag-of-frames approaches \cite{aucouturier2007bag}, which describes an auditory scene recording using summary statistics of short-time features. Unfortunately, this only captures the average structure of the scene, so the approach often fails when presented with highly dynamic scenes or those characterized by a few distinct sound events. Indeed, experiments in cognitive psychology \cite{dubois2006cognitive} and cognitive neuroscience \cite{nelken2004processing} suggest that this more closely matches human acoustic perception. We believe that the failure to model such distinct events is one of the reasons why this approach is insufficient for moderate-sized evaluation datasets \cite{lagrange:hal-01082501}.

Solving the acoustic similarity retrieval first requires the ability to capture meaningful signal structure at small time scales.
This is often achieved using mel-frequency cepstral coefficients (MFCCs).
Originally developed for speech processing \cite{davis-mermelstein}, MFCCs have recently found wider use in music information retrieval \cite{logan} and environmental audio processing \cite{aucouturier2007bag}.
A richer representation, the scattering transform, has enjoyed significant success in various audio \cite{Anden2014} and biomedical \cite{chudacek} signal classification tasks.
Its structure is that of a convolutional neural network \cite{lee,lostanlen-deep-spiral,soundnet,arandjelovic-zisserman}, but with fixed filters.
Specifically, it alternates convolutions with wavelet filters and pointwise nonlinearities to ensure time-shift invariance and time-warping stability \cite{Mallat2012}.

For our task, one advantage of the scattering transform is not requiring a training step, allowing for a wider range of applications compared to learned features. Indeed, for data mining of previously unheard datasets, the properties of relevant audio structures remains to be defined, leading to an unserpervised setting.

In this work, we propose a new model for acoustic scenes, where the signal is represented at sub-second scales by scattering transforms, while larger scales are captured by an cluster model. This unsupervised model quantizes the scattering coefficients into a given number of clusters. These clusters to define a set of distances for acoustic similarity retrieval. Evaluating this approach on a scene retrieval task, we obtain significant improvements over traditional bag-of-frames and summary statistics models applied both to MFCCs and scattering coefficients.

Motivations of the proposed approach and a brief review of the state of the art in acoustic scene modeling are given in Section \ref{sec:soa}. We describe the scattering transform in Section \ref{sec:scattering}, discuss feature post-processing in Section \ref{sec:design} and propose a cluster-based scene description in Section \ref{sec:object}. Section \ref{sec:experiments} describes several experiments for the acoustic scene similarity retrieval task. Results are reported in Section \ref{sec:results}.

@@

Computational bioacoustics refers to the numerical investigation of sound production, dispersion and reception in animals, including humans. A recent paradigm processes the audio stream in a holistic way over large time scales, without assuming that a single species is present throughout the recording. Automated systems within this paradigm attempt to infer global properties of bioacoustic scenes, including biodiversity indices \cite{Bardeli2010} and migration patterns \cite{Obrist2010}. They also mark time intervals of particular interest for detailed human inspection \cite{rosenstock2002landbird}.

A closely related field is that of urban sound environment analysis.
In this context, a popular approach is the bag of frames (BOF), first applied to the problem by Aucouturier et al. \cite{aucouturier2007bag}.
It models an auditory scene using high-level summary statistics computed from local features, typically implemented by Gaussian mixture models (GMMs) of MFCCs.
Recently, this representation was shown to perform comparably to direct averaging of the features for a variety of tasks \cite{lagrange:hal-01082501}.
This contrasts with the typical morphology of acoustic scenes, a ``skeleton of events on a bed of textures,'' where a few discrete sound events are superimposed upon a stationary acoustic background \cite{nelken2013}.
Such events are not well-characterized by summarizing short-term features, but are better described by large-scale temporal evolution of auditory scenes.
The latter approach should therefore prove more fruitful in measuring auditory scene similarity.

This statement has some support in auditory psychology as well as sound synthesis based on summary statistics \cite{mcdermott2013summary}. Studies in the cognitive psychology of urban sound environments have shown that global sound level (perceived or measured) is not sufficient to fully characterize an acoustic scene \cite{guyot2005urban,kang2006urban}. Instead, cognitive processes such as sound environment quality perception \cite{dubois2006cognitive} or loudness judgment \cite{kuwano_memory_2003} seem to rely upon higher-level cognitive attributes. These typically include the identities of the sound sources which consitute the scene. It has been shown that, if available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high-level cognitive classes. For example, in urban areas the presence of birds is likely to be heard in parks and are therefore strong pleasantness indicators. Consequently, research in sound perception is now strongly focused on the contribution of specific sound sources in the assessment of sound environments \cite{ricciardi2015sound,lavandier2006contribution}. Although the complete set of events occurring within a given auditory stream may not be discernable even to human expects, research has shown that a small set of events (so-called markers) suffice to reliably predict many high-level attributes.

From a cognitive psychology perspective, the consensus is therefore that only a few distinct events are sufficient to describe an auditory scene, in contrast to BOF models which treat each observation separately and do not capture their temporal structure. A method that takes this knowledge into account could therefore have potential for great impact in acoustic scene modeling, given a rich enough representation of these distinct events.

@@

Local invariance to time-shifting and stability to time-warping are necessary when representing acoustic scenes for similarity measurement.
The scattering transform is designed to satisfy these properties while retaining high discriminative power.
It is computed by applying auditory and modulation wavelet filter banks alternated with complex modulus nonlinearities.

\subsection{Invariance and stability in audio signals}
The notion of invariance to time-shifting plays an essential role in acoustic scene similarity retrieval.
Indeed, recordings may be shifted locally in time without affecting similarity to other recordings.
To discard this superfluous source of variability, signals are first mapped into a time-shift invariant feature space. These features are then used to calculate similarities. Since the features ensure invariance, it does not have to be learned when constructing the desired similarity measure.

Formally, given a signal $\boldsymbol{x}(t)$, we would like its translation $\boldsymbol{x_c}(t) = \boldsymbol{x}(t-c)$ to be mapped to the same feature vector provided that $|c| \ll T$ for some maximum duration $T$ that specifies the extent of the time-shifting invariance. We can also define more complicated transformations by letting $c$ vary with $t$. In this case, we have $\boldsymbol{x_\tau}(t) = \boldsymbol{x}(t-\tau(t))$ for some function $\tau$, which performs a time-warping of $\boldsymbol{x}(t)$ to obtain $\boldsymbol{x_\tau}(t)$. Time-warpings model various changes, such as small variations in pitch, reverberation, and rhythmic organization of events. These make up an important part of intra-class variability among natural sounds, so representations must be robust with respect to such transformations.

@@

The wavelet transform of an audio signal
$\boldsymbol{x}(t)$ is obtained by convolution with all wavelet filters.
Applying a pointwise complex modulus the transform yields
the wavelet scalogram
\begin{equation}
\boldsymbol{x_1}(t, \gamma_1)
= \vert \boldsymbol{x} \ast \boldsymbol{\psi_{\gamma_1}} \vert (t)\mbox{.}
\end{equation}
The scalogram bears resemblance to the constant-Q transform (CQT),
which is derived from the short-term Fourier transform (STFT) by averaging the frequency
axis into constant-Q subbands of center frequencies $2^{-\gamma_1/Q_1}\xi_1$.
Indeed, both time-frequency representations are indexed by time $t$ and log-frequency $\gamma_1$.
However, contrary to the CQT, the scalogram reaches a better time-frequency localization across the whole
frequency range, whereas the temporal resolution of the traditional CQT is fixed by the support of the STFT analyzing window. %\cite{Brown1992}
Therefore, the scalogram has a better temporal localization at high
frequencies than the CQT, at the expense of a greater computational cost
since the inverse fast Fourier transform routine must be called for each wavelet $\boldsymbol{\psi_{\gamma_1}}$ in the filter bank.
However, this allows us to observe amplitude modulations at fine temporal scales in the scalogram, down to $2Q_1/\xi_1$ for $\gamma_1 = 0$, of the order of $1\,\textrm{ms}$ given the aforementioned values of $Q_1$ and $\xi_1$.

To obtain the desired invariance and stability properties, the scalogram is averaged in time using a lowpass filter $\boldsymbol{\phi}(t)$ with cut-off frequency $1/T$ (and approximate duration $T$), to get
\begin{equation}
\mathbf{S_1}\boldsymbol{x}(t, \gamma_1) = \boldsymbol{x_1}(\cdot, \gamma_1) \ast \boldsymbol{\phi}(t),
\end{equation}
which is known as the set of first-order scattering coefficients. They capture the average spectral envelope of $\boldsymbol{x}(t)$ over scales of duration $T$ and where the spectral resolution varying with constant Q. In this way, they are closely related to the mel-frequency spectrogram and related features, such as MFCCs.

\subsection{Extracting modulations with second-order scattering}
In auditory scenes, short-time amplitude modulations may be caused by a variety of rapid mechanical interactions, including collision, friction, turbulent flow, and so on.
At longer time-scales, they also account for higher-level attributes of sound, such as prosody in speech or rhythm in music.
Although they are discarded while filtering $\boldsymbol{x_1}(t,\gamma_1)$ into the time-shift invariant representation $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$, they can be recovered from $\boldsymbol{x_1}(t,\gamma_1)$ by a second wavelet transform and another complex modulus.

We define second-order wavelets $\boldsymbol{\psi_{\gamma_2}}(t)$ in the same way as the first-order wavelets, but with parameters $\xi_2$, $J_2$, and $Q_2$. Consequently, they have center frequencies $2^{-\gamma_2/Q_2}\xi_2$ for $\gamma_2$ taking values between $0$ and $(J_2 Q_2 - 1)$. While this abuses notation slightly, the identity of the wavelets should be clear from context.
The amplitude modulation spectrum resulting from a wavelet modulus decomposition using these second-order wavelets is then
\begin{equation}
\boldsymbol{x_2}(t,\gamma_1,\gamma_2) =
\vert \boldsymbol{x_1} \ast \boldsymbol{\psi_{\gamma_2}} \vert(t,\gamma_1).
\end{equation}
In the following, we set $\xi_2$ to $2.5\,\mathrm{kHz}$, $Q_2$ to $1$, and $J_2$ to $12$. Lastly, the low-pass filter $\boldsymbol{\phi}(t)$ is applied to $\boldsymbol{x_2}(t, \gamma_1, \gamma_2)$ to guarantee local invariance to time-shifting, which yields the second-order scattering coefficients
\begin{equation}
\mathbf{S_2}\boldsymbol{x}(t,\gamma_1,\gamma_2) =
(\boldsymbol{x_2}(\cdot,\gamma_1,\gamma_2) \ast \boldsymbol{\phi})(t).
\end{equation}

The scattering transform $\mathbf{S}\boldsymbol{x}(t,\gamma)$ consists of the concatenation of first-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$ and second-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1,\gamma_2)$ into a feature matrix $\mathbf{S}\boldsymbol{x}(t,\gamma)$, where $\gamma$ denotes either $\gamma_1$ or $(\gamma_1,\gamma_2)$. While higher-order scattering coefficients can be calculated, for the purposes of our current work, the first and second order are sufficient.
Indeed, higher-order scattering coefficients have been shown to contain reduced energy and are therefore of limited use \cite{irene}.

@@

Given a set of $d$-dimensional feature vectors $X_u = \{x_1^u, \ldots, x_L^u\}$, extracted from the scene $s_u$, where $u=\lbrace 1,2,\ldots,U\rbrace$, we would like to partition $X_u$ into a set $C_u = \{c^u_1, \ldots, c^u_M\}$ of $M$ clusters. This partition is obtained by minimizing the variance of each cluster and known as a $k$-means clustering \cite{lloyd}.
Each scene $s_u$ is then described by a set of clusters $C_u$. Note that this quantization approach differs from unsupervised learning schemes such as the ones studied in \cite{bisot2016acoustic}, where the scene features are projected in a dictionary learned from the entire dataset. Here, with the aim of better balancing the influence of salient sound events and texture-like sounds on the final decision, the similarity between two scenes is computed based on the similarity of their centroids.

>>>>

\author{Vincent Lostanlen         \and
		Gr\'egoire Lafay  \and Joakim And\'en  \and Mathieu Lagrange.}

\institute{Vincent Lostanlen \at
              \email{vincent.lostanlen@nyu.edu}
\and
			Gr\'egoire Lafay \at
              \email{lafaygregoire@gmail.com}
\and
			Joakim And\'en \at
              \email{janden@flatironinstitute.org}           \and
			Mathieu Lagrange \at
              \email{mathieu.lagrange@cnrs.fr}
}

\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

@@

% As a general rule, do not put math, special symbols or citations
% in the abstract or keywords.
\begin{abstract}
The emerging field of computational ecoacoustics aims at retrieving high-level information from acoustic scenes recorded by some network of sensors. Large amounts of unlabeled data are gathered using such networks that needs to be analyzed. In order to identify which parts require human inspection, one needs tools that can mine data and identify structures such as recurring patterns or isolated events. To do so, one needs to be able to measure the level of similarity among the data without strong assumptions.

State of the art approaches for computing such a similarity is the ``bag-of-frames'' approach, i.e., modeling the audio signal using the summary statistics of short-term audio descriptors, such as mel-frequency cepstral coefficients (MFCCs).
Although they successfully characterize static scenes with little variability in auditory content, they are unable to accurately discriminate scenes with few salient events superimposed over a textured background.
To overcome this issue, we propose a two-scale representation which describes a signal at a short scale using scattering transforms while relying on a cluster model to characterize its larger-scale structure that better handle the required modeling of sparse events. Experiments performed using the acoustic scene similarity framework demonstrates the superiority of the proposed approach.
\keywords{
unsupervised learning \and data mining \and acoustic signal processing \and wavelet transforms \and audio databases \and content-based retrieval \and nearest neighbor searches \and acoustic sensors \and environmental sensors.}
% \PACS{PACS code1 \and PACS code2 \and more}
\end{abstract}

@@

The amount of audio data recorded from our sonic environment has grown considerably over the past decades.
In order to measure the effect of human activity and climate change on animal biodiversity, researchers have recently undertaken a massive deployment of acoustic sensors throughout the world \cite{warren2006urban, NessSST13, stowell13b}.
In addition, recent work has explored acoustic monitoring for characterization of human pleasantness in urban areas \cite{guyot2005urban, ricciardi2015sound}, as well as the prediction of annoyance due to traffic \cite{gloaguen}.
Since they bear a strong societal impact and raise many scientific challenges, we believe that these applications are of considerable interest for the signal processing community.

This work falls in the realm of ecoacoustics, an emerging field in computational bioacoustics, which is the interdisciplinary study of relationships between natural or anthropogenic sounds and their environments at the intersection of ecology, acoustics and computer science \cite{krause}. %Despite identified needs, the field of signal-based ecoacoustics is still in its infancy.
%Consequently, few well-designed datasets are readily available for evaluation purposes.
An important problem in this field is that manually analysing the recorded data to identify the quantities of interest is very costly \cite{wimmer2013sampling}. Some sort of pre-screening is therefore needed to reduce the need for human expert listening and annotation.

The most straightforward approach is to assume a closed set of events of interests like sounds of animals expected to live near the acoustic sensors, train models for them and perform automatic annotation \cite{}. A given time interval, for example a single day, is then represented by the number of events within the predefined typology that have been detected in that period of time. This approach allows the scientist to drastically reduce the amount of information that needs to be processed. However, this approach has two drawbacks. The first is the reliance to trained models whose behavior on unseen data (such as different sensors) is prone to errors and cannot be always be trusted. More importantly, it is based on \textit{a priori} knowledge and thus cannot be considered for exploratory analysis where the quantities of interest have yet to be defined.

In order to identify which parts needs human inspection, one needs tools that can mine data and annotate to whether recurring patterns  or sparsely distributed events to study anomalous or scarce behaviors. The latter allowing the scientist to then annotate only a representative audio of the periods of time with recurring patterns and the former to potentially discover unforseen phenomena.

With this aim, one needs to design an algorithm to perform an acoustic similarity retrieval, where the audio fragments judged ``most similar'' to a given query section are extracted from some larger dataset. This requires the audio fragments to be represented in some way that captures their distinctive qualities. This has been achieved by bag-of-frames approaches \cite{aucouturier2007bag}, which attempt to describe the auditory scene recording using summary statistics. Unfortunately, since this only captures the average structure of the scene, the approach often fails when presented with scenes which are dynamic and contain episodic, distinct sound events that discriminate the scene from others. Indeed, experiments in cognitive psychology \cite{dubois2006cognitive} and cognitive neurosciences \cite{nelken2004processing} suggest that this matches more closely human acoustic perception of acoustic scenes. We believe that the failure to model such distinct events is one of the reason why this approach is found to be insufficient when considering evaluation datasets of typical sizes \cite{lagrange:hal-01082501}.

%To design better representations for the similarity retrieval task, we draw inspiration from a closely related, and more mature, field of investigation: the classification of acoustic scenes. This task, consisting of predicting scene labels from audio recordings, is narrower in its range of applications, but has the advantage of being rooted in the cognitive psychology of categorization \cite{dubois2006cognitive,guastavino_ideal_2006}. In addition, it has received significant attention from the data mining community for some time, resulting in the availability of well-designed datasets and competitive algorithms for solving it.

Addressing this task first requires the ability to capture meaningful aspects within the signal at small time-scales.
In other words, how do we represent an extract of the auditory scene signal over a sub-second interval?
A standard tool for this are the mel-frequency cepstral coefficients (MFCCs), originally developed for speech processing \cite{davis-mermelstein}, but which have found wider use in recent years in music information retrieval \cite{logan} and environmental audio processing \cite{aucouturier2007bag}.
A richer set of features is found in the scattering transform, a generic signal representation which has enjoyed significant success in various audio classification tasks \cite{Anden2014,chudacek}.
This is because the scattering transform cascades wavelet convolutions and pointwise nonlinearities, thus imitating the architecture of a convolutional neural network \cite{lee,lostanlen-deep-spiral,soundnet,arandjelovic-zisserman}.

For the task at at hand, scattering features have the advantage of not requiring training and thus potentially lead to wider range of application than trained features. Indeed, the task at hand is the mining of yet unheard audio datasets and the properties of the audio that are of interest remains to be defined, thus leading to an unserpervised setting.

%The next question is the choice of a classifier. Here, a class of particularly successful algorithms train a classifier, such as a support vector machine (SVM) \cite{cristianini2000introduction}, to discriminate between short extracts of an auditory scene, and at the classification stage divide the signal into short extracts which are classified separately before pooling the classification results to reach a final decision. This last step is known as late integration and is crucial to the success of these methods, since it allows the classifier to concentrate on the short, distinct sound events that discriminate between scenes. The success of the late integration technique again leads credence to the idea that rare, distinct events in an auditory scene are sufficient to characterize it almost completely. Combining scattering transforms for low-level representation with late-integrated SVMs for classifying the scene at a higher level, we obtain a competitive auditory scene classification system, demonstrating the usefulness of both concepts for this task.

%we can emulate their selection behavior by quantizing the auditory scene and defining a distance over this quantization.
We also propose a new model for acoustic scenes, where the signal is represented at sub-second scales by scattering transforms, while larger scales are captured by an unsupervised method that quantizes the scattering coefficients into a given number of clusters. We then use these clusters to define a set of distances that can subsequently be used for similarity retrieval. Testing this approach on a scene retrieval task, we obtain significant improvements over traditional bag-of-frames and summary statistics models applied both to MFCCs and scattering coefficients.

Motivations of the proposed approach and a brief review of the state of the art in acoustic scene modeling are given in Section \ref{sec:soa}. We describe the scattering transform in Section \ref{sec:scattering}, discuss feature post-processing in Section \ref{sec:design} and propose a cluster-based scene description in Section \ref{sec:object}. Section \ref{sec:experiments} describes several experiments for the acoustic scene similarity retrieval task. Results are subsequently reported in Section \ref{sec:results}.

@@

%One avenue of ecoacoustics research, originating in bioacoustics, attempts to identify species properties of vocalizations, for example to link animal vocalization with behavior\cite{au2008principles, stowell2014large}.

Computational bioacoustics usually refers to the numerical investigation of sound production, dispersion and reception in animals (including humans). A recent paradigm  processes the audio stream in a holistic way over large time-scales, without assuming that a single species is present throughout the recording. Automated systems belonging to this paradigm attempt to infer global properties of bioacoustic scenes, including biodiversity indices \cite{Bardeli2010}, migration patterns \cite{Obrist2010}, as well as marking time intervals of particular interest for detailed human inspection \cite{rosenstock2002landbird}.

%The same division is found in the field of auditory scene analysis, wherein the interest may either be 1) to isolate and identify precise events, \ie acoustic event detection (AED), or 2) to predict global attributes assigned to large time spans of audio, \ie acoustic scene classification (ASC).
In the closely related field of urban sound environment analysis, a popular approach is the bag of frames (BOF) introduced in this context by Aucouturier et al. \cite{aucouturier2007bag}, where a scene is modeled by high-level summary statistics computed from local features.
Recently, the proposed implementation using Gaussian mixture models (GMMs) of mel-frequency cepstral coefficients (MFCCs) has been demonstrated to perform comparably to a direct averaging of the features for a variety of tasks \cite{lagrange:hal-01082501}.
This contrasts with the typical morphology of acoustic scenes, which we believe is a ``skeleton of events on a bed of textures'' \cite{nelken2013} where a few discrete sound events are superimposed upon a stationary acoustic background.
As those distinct and sparse events are often highly relevant for the task at hand, it is arguably more beneficial to focus on characterizing these distinct events through describing the large-scale temporal evolution of auditory scenes rather than on summary statistics of short-term features.

This statement has some support in auditory psychology as well as sound synthesis based on summary statistics \cite{mcdermott2013summary}. For example, studies in the cognitive psychology of urban sound environments have shown that global sound level (perceived or measured) is not sufficient to fully characterize an acoustic scene \cite{guyot2005urban,kang2006urban}. Rather, it seems that cognitive processes such as sound environment quality perception \cite{dubois2006cognitive} or loudness judgment \cite{kuwano_memory_2003} rely upon higher-level cognitive attributes such as the identities of sound sources which compose the scene. It has been shown that, if available, the complete description of the scene in terms of event occurrences is powerful enough to reliably predict high-level cognitive classes. For example, the presence of birds is very likely to be heard in parks in urban areas and so are strong pleasantness indicators. Consequently, research in sound perception is now strongly focused on the contribution of specific sound sources in the assessment of sound environments \cite{ricciardi2015sound,lavandier2006contribution}. Although the complete set of events occurring within a given auditory stream may not be discernable even to human expects, research has shown that a small set of events (so-called markers) suffice to reliably predict many high-level attributes.

From a cognitive psychology perspective, the consensus is therefore that only a few distinct events are sufficient to describe an auditory scene, in contrast to BOF approaches which treat each observation the same and do not incorporate temporal structure between features. A method that takes this knowledge into account could therefore have potential for great impact in the modeling of acoustic scenes, provided that the representation of the representation of those distinct events is rich enough.

%Aside from \cite{aucouturier2007bag} and \cite{lagrange:hal-01082501} that consider the task of acoustic scene similarity retrieval, much of the work dedicated to acoustic scene modeling is applied to the classification task.
%An overview of the efforts for the latter task is provided by the IEEE Audio and Acoustic Signal Processing (AASP) Detection and Classification of Acoustic Scenes and Events (DCASE) challenge \cite{barchiesi2015acoustic}, held in 2013 to compare different approaches to scene classification.

%The methods that performed the best all follow a similar structure, where low-level feature representations, such as MFCCs, are calculated over short windows and fed separately to a classifier, such as a support vector machine (SVM) \cite{cortes-vapnik,cristianini2000introduction}, for training. During the classification stage, a given recording is similarly processed into short windows represented by feature vectors which are classified individually by the classifier. The individial class assignments are then combined through a majority vote which assigns a class to the whole recording. This last step is known as late integration and allows the classifier to work on individual, shorter-scale sound events instead of trying to classify the recording as a whole. By doing this, the classifier is able to identify and exploit the distinct events that characterize the scene instead of trying to model its average structure. This is in agreement with the psychoacoustic results related to the perception of auditory scenes presented above.

%\ml{These two last paragraphs are a bit out of place here, since we don't do classification in the paper. Perhaps there is some way to cut them down a bit?}

%\gl{Othe this r paper:} \cite{jiang2005svm, kalinli2009saliency, su2011environmental, ye2015acoustic, bisot2015hog, chakrabarty2015exploring, xue2015auditory}

%\subsection{Datasets}

 % The latter is also considered in this paper as a validation of the proposed approach. \ja{What does this last sentence mean?}

%In this paper, we shall demonstrate that the use of features based on the wavelet scattering transform together with object-based representation of the acoustic scene allows us to achieve state-of-the-art \ja{?} performance in both the ASSR and ASC paradigms.

@@

Scattering transforms are time-shift invariant representations of audio signals computed by applying auditory and modulation wavelet filter banks alternated with complex modulus nonlinearities.
This section first explains the importance of local invariance to time-shifting and stability to time-warping in the representation of acoustic scenes, then describes how the scattering transform is designed to satisfy these properties while retaining high discriminative power.

\subsection{Invariance and stability in audio signals}
The notion of invariance to time-shifting plays an essential role in acoustic scene similarity retrieval as well as acoustic scene classification.
Indeed, recordings may be shifted locally without affecting their similarity to other recordings.
To discard this superfluous source of variability, signals are first mapped into a time-shifting invariant feature space. These features are then used to calculate similarities or to train a classifier. Since invariance is ensured by the features, it does not have to be learned during this last model construction step.

Formally, given a signal $\boldsymbol{x}(t)$, we would like its translation $\boldsymbol{x_c}(t) = \boldsymbol{x}(t-c)$ to be mapped to the same feature vector provided that $|c| \ll T$ for some maximum duration $T$ that specifies the extent of the time-shifting invariance. We can also define more complicated transformations by letting $c$ vary with $t$. In this case, we have $\boldsymbol{x_\tau}(t) = \boldsymbol{x}(t-\tau(t))$ for some function $\tau$, which performs a time-warping of $\boldsymbol{x}(t)$ to obtain $\boldsymbol{x_\tau}(t)$. Time-warpings model various changes, such as small variations in pitch, reverberation, and rhythmic organization of events. These make up an important part of intra-class variability among natural sounds, so it is important for representation to be robust to such transformations.

@@

The wavelet transform of an audio signal
$\boldsymbol{x}(t)$ is obtained by convolution with all wavelet filters.
Applying pointwise complex modulus the transform yields
the wavelet scalogram
\begin{equation}
\boldsymbol{x_1}(t, \gamma_1)
= \vert \boldsymbol{x} \ast \boldsymbol{\psi_{\gamma_1}} \vert (t)\mbox{.}
\end{equation}
The scalogram bears resemblance to the constant-Q transform (CQT),
which is derived from the short-term Fourier transform (STFT) by averaging the frequency
axis into constant-Q subbands of center frequencies $2^{-\gamma_1/Q_1}\xi_1$.
Indeed, both time-frequency representations are indexed by time $t$ and log-frequency $\gamma_1$.
However, contrary to the CQT, the scalogram reaches a better time-frequency localization across the whole
frequency range, whereas the temporal resolution of the traditional CQT is fixed by the support of the STFT analyzing window. %\cite{Brown1992}
Therefore, the scalogram has a better temporal localization at high
frequencies than the CQT, at the expense of a greater computational cost
since the inverse fast Fourier transform routine must be called for each wavelet $\boldsymbol{\psi_{\gamma_1}}$ in the filter bank.
However, this allows us to observe amplitude modulations at fine temporal scales in the scalogram, down to $2Q_1/\xi_1$ for $\gamma_1 = 0$, of the order of $1\,\textrm{ms}$ given the aforementioned values of $Q_1$ and $\xi_1$.

To obtain the desired invariance and stability properties, the scalogram is averaged in time using a lowpass filter $\phi(t)$ with cut-off frequency $1/T$ (and approximate duration $T$), to get
\begin{equation}
\mathbf{S_1}\boldsymbol{x}(t, \gamma_1) = \boldsymbol{x_1}(\cdot, \gamma_1) \ast \phi(t),
\end{equation}
which is known as the set of first-order scattering coefficients. They capture the average spectral envelope of $\boldsymbol{x}(t)$ over scales of duration $T$ and where the spectral resolution varying with constant Q. In this way, they are closely related to the mel-frequency spectrogram and related features, such as MFCCs.

\subsection{Extracting modulations with second-order scattering}
In auditory scenes, short-time amplitude modulations may be caused by a variety of rapid mechanical interactions, including collision, friction, turbulent flow, and so on.
At longer time-scales, they also account for higher-level attributes of sound, such as prosody in speech or rhythm in music.
Although they are discarded while filtering $\boldsymbol{x_1}(t,\gamma_1)$ into the time-shift invariant representation $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$, they can be recovered by a second wavelet transform and another complex modulus.

We define second-order wavelets $\boldsymbol{\psi_{\gamma_2}}(t)$ in the same way as the first-order wavelets, but with parameters $\xi_2$, $J_2$, and $Q_2$. Consequently, they have center frequencies $2^{-\gamma_2/Q_2}\xi_2$ for $\gamma_2$ taking values between $0$ and $(J_2 Q_2 - 1)$. While this abuses notation slightly, the identity of the wavelets should be clear from context.
The amplitude modulation spectrum resulting from a wavelet modulus decomposition using these second-order wavelets is then
\begin{equation}
\boldsymbol{x_2}(t,\gamma_1,\gamma_2) =
\vert \boldsymbol{x_1} \ast \boldsymbol{\psi_{\gamma_2}} \vert(t,\gamma_1).
\end{equation}
In the following, we set $\xi_2$ to $2.5\,\mathrm{kHz}$, $Q_2$ to $1$, and $J_2$ to $12$. Lastly, the low-pass filter $\phi(t)$ is applied to $\boldsymbol{x_2}(t, \gamma_1, \gamma_2)$ to guarantee local invariance to time-shifting, which yields the second-order scattering coefficients
\begin{equation}
\mathbf{S_2}\boldsymbol{x}(t,\gamma_1,\gamma_2) =
(\boldsymbol{x_2}(\cdot,\gamma_1,\gamma_2) \ast \phi)(t).
\end{equation}

The scattering transform $\mathbf{S}\boldsymbol{x}(t,\gamma)$ consists of the concatenation of first-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1)$ and second-order coefficients $\mathbf{S_1}\boldsymbol{x}(t,\gamma_1,\gamma_2)$ into a feature matrix $\mathbf{S}\boldsymbol{x}(t,\gamma)$, where $\gamma$ is a shorthand for either $\gamma_1$ or $(\gamma_1,\gamma_2)$. While higher-order scattering coefficients can be calculated, for the purposes of our current work, the first and second order are sufficient.
Indeed, higher-order scattering coefficients have been shown to contain reduced energy and are therefore of limited use \cite{irene}.

@@

Given a set of $d$-dimensional feature vectors $X_u = \{x_1^u, \ldots, x_L^u\}$, extracted from the scene $s_u$, where $u=\lbrace 1,2,\ldots,U\rbrace$, we would like to partition $X_u$ into a set $C_u = \{c^u_1, \ldots, c^u_M\}$ of $M$ clusters. The partitioning is done by minimizing squared error between the empirical mean, or centroid, of each cluster and the vectors belonging to it. This is known as $k$-means clustering \cite{lloyd} and has found widespread use for a variety of applications.
Each scene $s_u$ is then described by a set of clusters $C_u$. One should note that this quantization approach differs from unsupervised learning schemes such as the ones studied in \cite{bisot2016acoustic}, where the scene features are projected in a dictionary learned from the entire dataset. Here, with the aim of better balancing the influence of salient sound events and texture-like sounds on the final decision, the similarity between two scenes is computed based on the similarity of their centroids.
